My title

### 2024-03-16 paper/paper.qmd
Model updated

Based on our discussion in the model set-up section, the negative binomial distribution seems more suitable for our research objectives than the Poisson distribution. The Poisson model assumes that the mean and variance are equal, but in the real world, they are rarely equal. The negative binomial distribution incorporates a dispersion parameter, making the predictions more accurate.

@@

>>>>

Based on our discussion in the model set-up section, the negative binomial distribution seems more suitable for our research objectives than the Poisson distribution. The Poisson model assumes that the mean and variance are equal, but in the real world, they are rarely equal. As @tbl-mean-var shown below indicates, the mean and variance are not equal. The negative binomial distribution incorporates a dispersion parameter, making the predictions more accurate.

@@

We created a figure @fig-ppcheckandposteriorvsprior that includes both the Poisson and negative binomial models. For the Poisson model, the actual observed values y are represented in black. The predicted values $y_{Rep}$ are shown in light blue, which are replicated samples generated based on the posterior distribution of model parameters.

It can be seen that there is a significant overlap between many of the predictive distributions and the actual observed values, but at higher values, the range of the predictive distribution becomes very wide, indicating that the model may not fit well for high-count values.

For the negative binomial models, the actual observed values y are also represented in black, and the predicted values yrep are likewise shown in light blue, generated based on the posterior distribution of the model parameters. Compared to the Poisson model, the predictive distributions generated by the negative binomial model appear to be more concentrated, meaning they are more closely clustered around the actual observed values at higher counts, indicating that the negative binomial model may be more suitable for these data.
From the figure, it can be observed that the predictive distributions of the negative binomial model match the actual observed values better than those of the Poisson model. In the Poisson model's figure, the predicted values align well with the actual values near the peak in the middle but are wider in the tail distribution, which may mean that the Poisson model struggles with extreme values. In contrast, in the figure for the negative binomial model, the predicted values seem to follow the actual observed values more closely across the entire range, typically indicating it can better handle the overdispersion of the data.

To demonstrate that the negative binomial model is a better fit, we compared the Expected Log Predictive Density (ELPD) difference and the Standard Error (SE) difference between these two models. The ELPD difference column compares the difference in expected log pointwise predictive density between the two models—the negative binomial model and the Poisson model. ELPD is an indicator of model predictive performance, with higher ELPD values typically indicating better model predictive performance. For the negative binomial model (listed as cause_of_death_alberta_neg_binomial), the ELPD difference is 0.0. For the Poisson model (listed as cause_of_death_alberta_poisson), the ELPD difference is -6160.6, meaning that, relative to the negative binomial model, the Poisson model's predictive performance is worse.

The SE difference represents the uncertainty in the estimation of the ELPD difference. A smaller SE difference indicates that the estimation of the ELPD difference is more precise; a larger SE difference indicates that the estimation of the ELPD difference is less stable and has more uncertainty. In other words, it tells us the reliability of the difference in predictive performance between the models. The SE difference is 1412.1, which is a relatively large value, indicating considerable uncertainty in the estimation of the predictive performance difference between the negative binomial model and the Poisson model. Nonetheless, the significant difference in ELPD (-6160.6) compared to its SE suggests that this difference is statistically significant. In other words, even considering the uncertainty in the ELPD difference, the negative binomial model's predictive performance is significantly better than that of the Poisson model, supporting our previous statements.
